\vspace{-2cm}A generative model for urban firm networks

### 2020-08-31 main.tex
Update on Overleaf.

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{natbib,graphicx,mdframed,hyperref,threeparttable,amssymb}

\title{\vspace{-2cm}A generative model for urban firm networks}

\date{}

\begin{document}

\author{J. Raimbault$^{1,2,3,\ast}$, N. Zdanowska$^{1,3}$\smallskip\\
$^{1}$ Centre for Advanced Spatial Analysis, UCL\\
$^{2}$ UPS CNRS 3611 ISC-PIF\\
$^{3}$ UMR CNRS 8504 G{\'e}ographie-cit{\'e}s\\
$^{\ast}$ \texttt{juste.raimbault@polytechnique.edu}
}

\maketitle
% NetSci format: A4, 11pt, 1in margin, max 12p + up to 3p references

\begin{abstract}
The emergence of entangled urban networks is a salient feature of globalisation processes. Understanding the drivers behind the growth of such networks, and in particular urban firm networks, is crucial for the economic resilience of urban systems. We introduce in this paper a generative network model for firm networks at the urban area level, which includes several complementary processes which are the economic size of urban areas at origin and destination, industrial sector proximity between firms, the strength of links from the past as well as the geographical and socio-cultural distance. An empirical network analysis on European firm ownership data confirms a role for each factor. We then simulate network growth for synthetic systems of cities, unveiling stylized facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range. We finally calibrate the model on the European network, outperforming statistical models and showing a strong role of path-dependency. Potential applications of the model include the study of mitigation policies for exogenous shocks such as economic crisis.

\end{abstract}

\section{Introduction}

The intensification of globalisation of the economy since the late 20th century resulted in the rise of the network society \cite{castells2000networksociety} and the emergence of a world-city system \cite{taylor2001specification}, characterized by highly interconnected urban centres at the global level \cite{sassen1991}. These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity implying spatial and non spatial patterns \cite{martinus2018global}. Interactions of economic nature between transnational firms, are together with trade among the most representative of the current economic trends \cite{taylor2001specification}. Understanding the drivers of growth and geographical properties of such a urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}. Straightening these city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}, by developing strategic industries for global insertion of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}. The functioning of such a city-network and the position of cities within these networks \cite{gluckler2016relational} can also permit to infer the consequences of future exogenous shocks such as an economic crisis, a single market collapse or a post-Brexit scenario for the European Union.

The use of network models to understand processes underlying the emergence and dynamics of such urban networks has already been widely tackled in the literature. \cite{dai2016simulating} propose a generative urban network model at the regional scale, combining geographical factors with network topological properties. Air transportation networks can also be understood as urban networks, and \cite{zanin2013modelling} reviews several models for such networks. Trade network have been extensively studied from complex networks perspective, but are generally driven at the country-level \cite{garlaschelli2005structure} because of lack of city-level data. Input-output models, mostly used in regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}. In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for abstract flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2018indirect}.

In this paper we choose to focus on urban networks of firms to question how can we understand and capture geographical and economic processes as agglomeration economies \cite{Glaeser2001} characteristic to spatial economy \cite{Fujitaetal2001} within a generative model. The specific question we tackle is the complementarity of different processes to generate such urban networks. One originality of the present paper is to approach these issues within the system of cities framework \cite{berry1964cities}, where the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}. We examine the interactions of European cities within firm linkages defined by corporate ownership links. Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geographical structures \cite{2019arXiv191014652Z}.

More precisely, our contribution relies on the following points: (i) we propose an empirical analysis of the firm ownership urban network for the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database; (ii) we introduce a generative network model to simulate the growth of such linkages at urban areas scale, which combines multiple factors influencing link formation, as economic intensity of origin and destination of urban areas, industrial sector proximity, the strength of previous links, and geographical and socio-cultural distance; (iii) the model allows us to compare the effect of different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; and (iv) we calibrate the model on the empirical network.

The rest of this paper is organized as follows: we first proceed to an empirical analysis of the network studied. We then describe the generative model for urban networks, and its sensitivity analysis and exploration on synthetic data. We then parametrize and calibrate the model on real European firm ownership data. We finally discuss implications of our results and possible extensions.

\section{Empirical network analysis}

Companies from the AMADEUS database are georeferenced using the Geonames database (using postcode or address depending on the information available). We then join this data with the GHS-FUA dataset for Functional Urban Areas \cite{Florczyk2019ghs}, which are the most relevant objects for applying our model. Starting from a firm-level dataset of 1,562,862 nodes and  1,866,936 links, we obtain a directed network of 573 urban areas and 9323 ownership links. Weight of links are obtained by computing the owned share of turnover at destination, i.e. $w_{ij} = \sum_{k \in i,l \in j} p_{kl} \cdot T_l$ where $p_{kl}$ of the ownership share of company $l$ by company $k$ and $T_l$ the turnover of company $l$. Node attributes include an economic size of urban areas, that we compute as the sum of turnovers of companies within the area. This quantity is highly correlated with area-level GDP included in the GHS database (Spearman correlation $\rho = 0.71$) and also with population ($\rho = 0.60$), so we expect to capture size effects while keeping the consistence of a single main data source. An other attribute of nodes is the industrial sector composition, expressed as a proportion of turnover in the area associated to a given industrial sector. Detailed sectors of firms are available up to the four digits of the NACE classification in the raw data. We consider the first level classification (21 categories) and compute sector proportions accordingly. This allows us to define an industrial proximity matrix between urban areas, taking a cosine similarity between vectors as $s_{ij} = \sum_{k=1}^{K} S_{ik}\cdot S_{jk}$ where $S_{ik}$ are sector proportions.

The urban network has heavy tail weighted degree and edge weight distributions, as shown in Fig.~\ref{fig:nwdist}. We fit power law and log-normal distributions, including minimal value cutoff following \cite{clauset2009power}, using the \texttt{powerlaw} R package \cite{powerlawpackage}. For both, log-normal distributions appear to be a better fit ($\mu=18.8$, $\sigma=2.3$ for weighted degree; $\mu=13.8$, $\sigma=2.8$ for edge weight), and include a much broader range of the empirical distributions with lower estimated cut-offs. These heavy tail properties suggest the relevance of a generative model including self-reinforcing processes which are known to produce such distributions.

Regarding the community structure of the network, different modularity maximisation algorithms (run on the undirected corresponding network) give a same undirected modularity of 0.38, and a directed modularity \cite{nicosia2009extending} of 0.34 for the greedy algorithm of \cite{clauset2004finding} and 0.36 for the Louvain algorithm \cite{blondel2008fast}, corresponding to 36 communities (resp. 15). In comparison, with 29 countries, the classification of urban areas by country has a modularity of 0.32. This means that the structure of flows present certain regional patterns. Indeed, a null model assigning random labels with 30 communities, bootstrap 1000 times, yields an average modularity of $0.049 \pm 0.002$, confirming the statistical significance of these communities.

To have a first insight into factors playing a role in the formation of links, we proceed to a statistical analysis similar to unconstrained spatial interaction models \cite{wilson1975some}. We test five different statistical models, including progressively the processes candidate to be included in the network generative model. The most general statistical model is
\[
\log(w_{ij}) = \log(d_{ij}) + \log(T_i) + \log(T_j) + \log(s_{ij}) + \alpha_{c_i,c_j} + \varepsilon_{ij}
\]
where $\alpha_{c_i,c_j}$ is a fixed effect term for the couple of countries $c_i,c_j$. We use a multiplicative form by taking logarithms as it is usually done for spatial interaction models. Furthermore, origin and/or destination constrained models are not relevant in our case as links include not only turnover but also a share of ownership. Estimation results are given in Table~\ref{tab:reg}. We also fit a Poisson generalized linear model with a logarithmic link on the rounded weights \cite{flowerdew1988fitting}. We find that distance alone (model 1) has a significant negative effect but with an explained variance close to zero. The second model adds fixed effects by pair of countries, for which around a third are significant, and with a better $R^2$ of 0.17. Adding origin and destination economic size (turnovers $T_i$ and $T$, model 3) slightly improves the explanative power, but the best model among the linear ones is the model 4 with all factors and fixed effects, both in terms of explained variance ($R^2 = 0.31$) and Akaike information criterion. The latter indicates the absence of overfitting. The Poisson model with all variables is much better regarding pseudo-$R^2$ and provides much more significant estimates (all fixed effects significant and much smaller standard deviations). Across all models, we find consistent qualitative stylized facts: (i) distance has a negative influence; (ii) economic sizes have a positive influence, with the size of the origin being more important (what is consistent with ownership links, strongly driven by decision at the headquarters); (iii) industrial proximity has a positive influence, i.e. similarity is more important than complementarity for this dimension; (iv) an important part of country fixed-effects are significant; and (v) the order of magnitude of the influence of each factor is roughly the same, in the sense that no process totally dominates the others. This suggests the relevance of including all these different dimensions in the generative network model.

%%%%%%%%%%%%%
\begin{table}[h]
\caption{Ordinary Least Square estimations of statistical models explaining $\log(w_{ij})$ (standard errors in parenthesis). ***, ** and * respectively indicate 0.01, 0.05 and 0.1 levels of significance. For the fixed effect by origin-destination pairs of countries, we give the proportion of significant coefficients (where $p<0.1$). Models (1) to (4) are ordinary least square models, while model (5) is a Poisson generalized linear model with logarithmic link. We compare model performance using $R^2$, Mean Square Error, and Akaike Information Criterion. For the Poisson model, AIC is not given as it is not comparable, and the pseudo-$R^2$ is given either.\label{tab:reg}}
\medskip
\begin{tabular}{|l|c|c|c|c|c|}
\hline
Model  & (1) & (2) & (3) & (4) & (5) \\ 
\hline
$\log(d_{ij})$ &      -0.06** (0.03) &   -0.11** (0.05)  & -0.41*** (0.02)  & -0.35*** (0.04)  &  -0.26*** ($5\cdot 10^{-6}$) \\
$\log(T_i)$ &   &   & 0.56*** (0.01) &  0.56*** (0.01) & 0.79*** ($2\cdot 10^{-6}$) \\
$\log(T_j)$ &     &   & 0.39*** (0.01) &  0.39*** (0.01) & 0.66***  ($1.5\cdot 10^{-6}$) \\
$\log(s_{ij})$ &     &   &  &  0.19*** (0.07) & 0.53*** ($9\cdot 10^{-6}$)  \\
Countries &    &  28.3\% &   &  26.5\% & 100\% \\
\hline
$R^2$ &       0.00059   &  0.17 & 0.21 &  0.31  &  0.56 \\
MSE & 7.75 & 6.84 & 6.10 & 5.33 & 8.72 \\
AIC &        44304   &  43578  &  42131  & 41917 &   \\
\hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%

\section{Model}

\subsection{Rationale}

The model is expected to capture a single macroscopic urban scale, i.e. the level of the urban system where basic entities are cities. Links are induced by underlying firms but these are not explicit. Cities are defined by their economic size, but also by an industrial sector profile. As \cite{martinus2018global} point out, a combination of several distances may play a role in establishing linkages. Geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}. Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalized to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets despite the European common market (confirmed relevant by the fixed effect regression described above); (iii) city economic size, which has furthermore scaling laws properties; (iv) industrial proximity in terms of firm sector composition; (v) previous linkages (influence of history). This last point is crucial to capture accumulated effects and include path-dependency, which fits well the use of a simulation model instead of a more simple statistical model.

\subsection{Model description}

@@

\begin{equation}
p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_F} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_T} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s\left(S_{ik},S_{jk}\right)^{\gamma_S} \cdot \exp \left(- \gamma_G \cdot d_{ij}\right) \cdot \exp \left(- \gamma_D \cdot g_{ij}\right)
\end{equation}

@@

Note that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalize the exponents to 1 (which is different than renormalising the probability), to ensure to include convex functions. The ``geopolitical/sociocultural'' distance remains abstract and should be parametrised (see setup below) and estimated in the data-driven setting. We assume no evolution of economic size in this simple version of the model, which means that adjusted parameter are valid on a restricted time period.

\subsection{Model setup}

@@

The synthetic system of cities is generated to be at the scale of a continent with properties similar to Europe: (i) $N=700$ cities with an empirical power-law from the Global Human Settlements database for economic sizes ($\alpha = 1.1$), distributed randomly into a uniform and isotropic square space of width $d_{max}=3000km$; (ii) clustered into $C = 20$ countries using a k-means algorithm (this remains consistent with non-correlated city sizes at the scale of the countries as our cities are randomly distributed in space \citep{simini2019testing}); (iii) with synthetic distribution of sectors following log-normal distributions adjusted in a way that large cities have more high-value activities and are more diverse. More precisely for this last point, assuming sectors are distributed continuously along a one-dimensional axis, cities are assumed to have a sector distribution with a log-normal density function. The largest city is constrained to have a standard deviation and a mode of $1/2$ and the smallest $1/K$, with a linear interpolation between the two for other cities. Parameters of the log-normal law are obtained by solving the corresponding two equations system for each city, and each distribution is discretised and renormalised to be a vector of probabilities. In synthetic experiments, the network is generated from an empty initial network until reaching $t_f=1500$, with a unit link volume of $w_0 = 1$.

@@

The real setup is done using the empirical network analyzed above. Urban areas are distributed according to their positions in the actual geographical space, and economic size and sector composition attributes are used. The socio-economic distance is constructed using information from the statistical modeling, by considering fixed-effect coefficients estimated for each pair of countries. Initial network is empty, and the real links are taken as targets for the generated network. We constrain the total volume of flows by setting $w_0 = \sum_{ij} w_{ij} / t_{f}$ when the number of time steps $t_f$ is given as a parameter.

\subsection{Model parameters and indicators}

The model parameters are the Cobb-Douglas weights $\gamma_F,\gamma_T,\gamma_W,\gamma_s$ and distance decay parameters $\gamma_G,\gamma_D$. The model also includes path-dependency by integrating the influence of previous links. Therefore it can not be reduced to a simple statistical model. Finally, the role of space introduces even further complexity, yielding a generative model, which has to be simulated.

We consider several macroscopic indicators to quantify the generated network. Firstly, geographical structures are captured by (i) internationalisation (modularity of countries in the network); (ii) metropolisation (correlation between weighted degree and city size). Two other thematic indicators, which are regionalisation (correlation between length and flow of links, stratified by size of extremities), and specialisation (correlation between sector proximity and flow of links, stratified by size of extremities), are not included in this study for the sake of simplicity as each has a high dimension. We also consider several network indicators as Louvain modularity, sizes of maximal modularity communities, degree and link weight distributions (described by their average, hierarchy, and entropy), and correlations between degree and city size, and between link weight and distance.

\section{Model exploration and calibration}

@@

\subsection{Synthetic city systems}

@@

%%%%%%%%%%%%%%%
\section{Discussion}

>>>>

\documentclass[10pt,letterpaper]{article}
\input{header.tex}

\usepackage{ragged2e}
\begin{document}
\justify
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{A generative model for urban firm networks} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
% alternative titles?
% Modeling the growth of urban firm networks
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Juste Raimbault\textsuperscript{1,2,3*},
Natalia Zdanowska\textsuperscript{1,3},
Elsa Arcaute\textsuperscript{3}
\\
\bigskip
\textbf{1} Centre for Advanced Spatial Analysis, UCL, London, United Kingdom
\\
\textbf{2} UPS CNRS 3611 ISC-PIF, Paris, France
\\
\textbf{3} UMR CNRS 8504 G{\'e}ographie-cit{\'e}s, Paris, France
\\
\bigskip

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* juste.raimbault@polytechnique.edu

\end{flushleft}

% Plos one requirements https://journals.plos.org/plosone/s/submission-guidelines
% Deadline for the Cities as Complex Systems Call for Papers: 4th September https://collections.plos.org/s/cities
% Manuscripts can be any length. There are no restrictions on word count, number of figures, or amount of supporting information. 
% PLOS uses “Vancouver” references style
% We need to reorganise the paper to have more or less the following sections (which can be reordered or renamed):
% 1. Title page: List title, authors, and affiliations as first page of manuscript 2. Abstract 3. Introduction 4. Materials and Methods 5. Results 6. Discussion 7. Conclusions (optional)

\section*{Abstract}
The emergence of interconnected urban networks is a crucial feature of globalisation processes. Understanding the drivers behind the growth of such networks -- in particular urban firm networks --, is essential for the economic resilience of urban systems. We introduce in this paper a generative network model for firm networks at urban area level including several complementary processes: the economic size of urban areas at origin and destination, industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distance. An empirical network analysis on European firm ownership data confirms the relevance of each of these factors. We then simulate network growth for synthetic systems of cities, unveiling stylized facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range. We finally calibrate the model on the European network, outperforming statistical models and showing a strong role of path-dependency. Potential applications of the model include the study of mitigation policies to deal with exogenous shocks such as economic crisis or potential lockdowns of countries.

\linenumbers

\section*{Introduction}

\cite{derudder2018central}
\cite{storme2019introducing} cities firms interactions

Networks constitute the new social morphology of our societies, since the intensification of globalisation of the economy in the late 20th century \cite{castells2000networksociety}. A world-city system has since then emerged \cite{taylor2001specification}, characterized by highly interconnected urban centres at the global level \cite{sassen1991}. These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity implying spatial and non-spatial patterns \cite{martinus2018global}. Interactions of economic nature between transnational firms, are together with international trade patterns among the most representative of the current economic trends \cite{taylor2001specification}. Understanding the drivers of growth and geographical properties of such an urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}. Straightening these city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}, by developing strategic industries for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}. The functioning of such a city-network and the position of cities within these networks \cite{gluckler2016relational} can also permit to infer the consequences of future exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.

The use of network models to understand processes underlying the emergence and dynamics of such networks has already been tackled in the literature, but still not extensively at city level. Generative urban network models combining geographical factors with network topological properties exist at the regional scale {\cite{dai2016simulating}}. %Air transportation networks can also be understood as urban networks. \cite{zanin2013modelling} reviews several models for such networks.
Trade networks have been extensively studied from the complex networks perspective, but these are generally driven at the country-level \cite{garlaschelli2005structure} because of a lack of city-level data. Input-output models, mostly used in regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}. In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for different types of flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2018indirect}.

In this paper we choose to focus on urban networks of firms to question and capture geographical and economic processes as internationalisation, metropolisation, regionalisation and specialisation of cities within a generative model. The specific question we tackle is the complementarity of different processes to generate such urban networks. The originality of the present paper is to examine these issues within the system of cities framework \cite{berry1964cities}. In this approach the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}. We examine the interactions of European cities within firm linkages defined by corporate ownership links. Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2019arXiv191014652Z}.

More precisely, our contribution consists in the following: (i) we propose an empirical analysis of the firm ownership urban network in the European Union, based on the Bureau Van Dijk's \emph {AMADEUS} database \cite{AMADEUS2018}; (ii) we introduce a generative network model to simulate the growth of such linkages at urban areas scale, which combines multiple factors influencing link formation, as economic intensity of origin and destination of urban areas, industrial sector proximity, the strength of previous links, and geographical and socio-cultural distance; (iii) the model allows us to compare the effect of different factors on the final network structure, which is extensively studied through model sensitivity analysis and exploration; and (iv) we calibrate the model on the empirical network.

The rest of this paper is organized as follows: We first describe the generative model for urban networks and proceed to an empirical analysis of the network used as a case study. We then do a sensitivity analysis and exploration of the model on synthetic data. The model is calibrated on real European firm ownership data. We finally discuss implications of our results and possible extensions.

\section*{Materials and Methods}

\subsection*{Generative urban network model}

\subsubsection*{Rationale}

The model is expected to capture a single macroscopic urban scale, i.e. the level of the urban system where basic entities are cities. Links are induced by underlying firms but these are not explicit. Cities are defined by their economic size, but also by an industrial sector profile. As the authors of \cite{martinus2018global} point out, a combination of several distances may play a role in establishing linkages. Geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}. Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalized to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets within the European common market (confirmed relevant by the fixed effect regression described above); (iii) city economic size, which has furthermore scaling laws properties; (iv) industrial proximity in terms of firm sector composition; (v) previous linkages (influence of already existing links in the past). This last point is crucial to capture accumulated effects and include path-dependency, which fits well the use of a simulation model instead of a more simple statistical model.

\subsubsection*{Model description}

@@

\begin{equation}
p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_O} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_D} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s\left(S_{ik},S_{jk}\right)^{\gamma_S} \cdot \exp \left(- d_{ij} / d_0\right) \cdot \exp \left(- c_{ij} / c_0\right)
\end{equation}

@@

Note that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalise the exponents to 1 (which is different than renormalising the probability), to ensure to include convex functions. The ``geopolitical/sociocultural'' distance remains abstract and should be parametrised (see setup below) and estimated in the data-driven setting. We assume no evolution of economic size in this simple version of the model, which means that adjusted parameter are valid on a restricted time period.

\subsubsection*{Model setup}

@@

The synthetic system of cities is generated to be at the scale of a continent with properties similar to Europe: (i) $N=700$ cities with an empirical power-law from the Global Human Settlements database for economic sizes ($\alpha = 1.1$), distributed randomly into a uniform and isotropic square space of width $d_{max}=3000km$;
(ii) clustered into $C = 20$ countries using a k-means algorithm (this remains consistent with non-correlated city sizes at the scale of the countries as our cities are randomly distributed in space \cite{simini2019testing});
(iii) with synthetic distribution of sectors following log-normal distributions adjusted in a way that large cities have more high-value activities and are more diverse.

More precisely for this last point, assuming sectors are distributed continuously along a one-dimensional axis, cities are assumed to have a sector distribution with a log-normal density function. The largest city is constrained to have a standard deviation and a mode of $1/2$ and the smallest $1/K$, with a linear interpolation between the two for other cities. Parameters of the log-normal law are obtained by solving the corresponding two equations system for each city, and each distribution is discretised and renormalised to be a vector of probabilities. See \ref{S2_Text} for all details of this synthetic setup.

In synthetic experiments, the network is generated from an empty initial network until reaching $t_f=1500$, with a unit link volume of $w_0 = 1$.
%\todo{explore effect of $t_f$ and/or $w_0$; be closer to real setup?} % -> more interesting to explore effect of urban hierarchy

@@

The real setup is done using the empirical network analyzed above. Urban areas are distributed according to their positions in the actual geographical space, and economic size and sector composition attributes are used. The socio-economic distance is constructed using information from the statistical modeling, by considering fixed-effect coefficients estimated for each pair of countries. See \ref{S2_Text} for more details on the construction of the socio-economic distance matrix.

Initial network is empty, and the real links are taken as targets for the generated network. We constrain the total volume of flows by setting $w_0 = \sum_{ij} w_{ij}^{(obs)} / t_{f}$ when the number of time steps $t_f$ is given as a parameter.

\subsection*{Model parameters and indicators}

The model parameters are the Cobb-Douglas weights $\gamma_F,\gamma_T,\gamma_W,\gamma_s$ and distance decay parameters $\gamma_G,\gamma_D$. As the model also includes path-dependency by integrating the influence of previous links, it can not be reduced to a simple statistical model. Finally, the role of space introduces even further complexity, yielding a generative model, which has to be simulated.

We consider two main macroscopic indicators to quantify the generated network. Firstly geographical structures are captured by internationalisation (modularity of countries in the network). The situation of a minimised modularity corresponds to dense connections between cities of different countries of Europe and sparse connections between cities of the same European country. In practice, modularity is computed by
\begin{equation}

\end{equation}
% put the actual modularity
% add some reference values - some kind of null model

Secondly we consider the metropolisation indicator as the correlation between the weighted degree and the GDP of the city. With this indicator we capture the effect of large cities or metropolises of regional importance attracting the biggest amount of links and involved in high interactions with other cities. It is computed with a Pearson correlation estimator as
\begin{equation}
    M = \hat{\rho} \left[d_i , E_i \right]
\end{equation}

%I think we need to keep this paragraph but shorten it, because to my understanding network average community size is tested in Figure 5, but not the Louvain modularity for ex.
%We also consider several network indicators as Louvain modularity, sizes of maximal modularity communities, degree and link weight distributions (described by their average, hierarchy, and entropy), and correlations between degree and city size, and between link weight and distance.

We take a geographical approach to study the behavior of the generative model. Other related approaches include indicators of economic complexity \cite{hidalgo2007product}. % develop
These are less relevant in our case as the sector structure of cities does not evolve in the model.

\subsection*{Empirical data}

Companies from \emph {AMADEUS} are georeferenced using the Geonames database (using postcode or address depending on the information available). We then join this data with the Global Human Settlement dataset GHS-FUA for Functional Urban Areas \cite{Florczyk2019ghs} -- the most relevant objects for applying our model. Starting from a firm-level dataset of 1,562,862 nodes and 1,866,936 links, we obtain a directed network of 573 urban areas and 9323 ownership links. The weight of links is obtained by computing the owned share of turnover at destination, i.e. $w_{ij} = \sum_{k \in i,l \in j} p_{kl} \cdot T_l$ where $p_{kl}$ of the ownership share of company $l$ by company $k$ and $T_l$ the turnover of company $l$. Node attributes include an economic size of urban areas, that we compute as the sum of turnovers of companies within the area. This quantity is highly correlated with the GDP of the area-level included in the GHS database (Spearman correlation $\rho = 0.71$) and also with population ($\rho = 0.60$). We then expect to capture size effects while keeping the consistence of a single main data source. An other attribute of the nodes is the industrial sector composition, expressed as a proportion of turnover in the area associated to a given industrial sector. Detailed sectors of firms are available up to the four digits of the NACE classification in the raw data \cite{EUROSTAT2008}. We consider the first level classification (21 categories) and compute sector proportions accordingly. This allows us to define an industrial proximity matrix between urban areas, taking a cosine similarity between vectors as $s_{ij} = \sum_{k=1}^{K} S_{ik}\cdot S_{jk}$ where $S_{ik}$ are sector proportions.

The urban network has heavy tail weighted degree and edge weight distributions, as shown in Fig.~\ref{fig:nwdist}. We fit power law and log-normal distributions, including minimal value cutoff following \cite{clauset2009power}, using the \texttt{powerlaw} R package \cite{powerlawpackage}. For both, log-normal distributions appear to be a better fit ($\mu=18.8$, $\sigma=2.3$ for weighted degree; $\mu=13.8$, $\sigma=2.8$ for edge weight), and include a much broader range of the empirical distributions with lower estimated cut-offs. These heavy tail properties suggest the relevance of a generative model including self-reinforcing processes, which are known to produce such distributions.

Regarding the community structure of the network, different modularity maximisation algorithms (run on the undirected corresponding network) give the same undirected modularity of 0.38, and a directed modularity \cite{nicosia2009extending} of 0.34 for the greedy algorithm of \cite{clauset2004finding} and 0.36 for the Louvain algorithm \cite{blondel2008fast}, corresponding to 36 communities (resp. 15). In comparison, with 29 countries, the classification of urban areas by country has a modularity of 0.32. This means that the structure of flows present certain regional patterns. Indeed, a null model assigning random labels with 30 communities, bootstrap 1000 times, yields an average modularity of $0.049 \pm 0.002$, confirming the statistical significance of these communities. We show in Fig.~\ref{fig:map_communities} a map of the communities obtained.

%%%%%%%%%%%%%%
\begin{figure}
    \centering
    \includegraphics{}
    \caption{{\bf Map of network communities obtained by modularity maximisation.}}
    \label{fig:map_communities}
\end{figure}
%%%%%%%%%%%%%%

\subsubsection*{Statistical analysis}

In order to understand which are the most determining geo-economic factors influencing creation of links between cities, we run a statistical analysis as applied in the case of unconstrained spatial interaction models \cite{wilson1975some}. We test five different statistical models, including progressively the processes to be included in the network generative model. The most general statistical model is:

\begin{equation}
\log(w_{ij}) = \log(d_{ij}) + \log(T_i) + \log(T_j) + \log(s_{ij}) + \alpha_{c_i,c_j} + \varepsilon_{ij}
\end{equation}

where $\alpha_{c_i,c_j}$ is a fixed effect term for the coupling of countries $c_i,c_j$. A multiplicative form is applied by taking logarithms as usually done for spatial interaction models. Furthermore, origin and/or destination constrained models are not relevant in our case, as links include information about the turnover and a share of ownership at the same time. Estimation results are given in Table~\ref{tab:reg}. We also fit a Poisson generalized linear model with a logarithmic link on the rounded weights \cite{flowerdew1988fitting}. The findings show that when distance only is taken into account (model 1), it has a significant negative effect on link creation, but with an explained variance close to zero. Model 2 adds fixed effects by pair of countries, for which around a third are significant, and with a better $R^2$ of 0.17. Adding the economic size at origin and destination (turnovers $T_i$ and $T$, model 3) slightly improves the explanative power (model 3). The best model among the linear ones is the model 4, with all factors and fixed effects, both in terms of explained variance ($R^2 = 0.31$) and Akaike information criterion. Model 4 also indicates the absence of overfitting. Model 5, which is a Poisson model with all variables is much a better fit regarding pseudo-$R^2$ and provides much more significant estimates (all fixed effects significant and much smaller standard deviations). Across all models, consistent qualitative stylized facts regarding link creation between cities are revealed: (i) distance has a negative influence; (ii) economic sizes have a positive effect, with the size of the origin city being more important (consistent in the ownership links context, as strongly driven by decision at the headquarter/owner level); (iii) industrial proximity has a positive influence, i.e. similarity is more important than complementarity for this dimension; (iv) an large part of country fixed-effects are statistically significant; and (v) the order of magnitude of the influence of each factor is roughly the same, which means that no process totally dominates the others.
This confirms the relevance of including all these different dimensions in the generative network model, since they have a comparable explanatory role in statistical models.

%%%%%%%%%%%%%
\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in}
\caption{{\bf Statistical models.} Ordinary Least Square estimations of statistical models explaining $\log(w_{ij})$ (standard errors in parenthesis). ***, ** and * respectively indicate 0.01, 0.05 and 0.1 levels of significance. For the fixed effect by origin-destination pairs of countries, the proportion of significant coefficients is attributed (where $p<0.1$). Models (1) to (4) are ordinary least square models, while model (5) is a Poisson generalized linear model with logarithmic link. We compare model performance using $R^2$, Mean Square Error, and Akaike Information Criterion. For the Poisson model (model 5), AIC is not given as it is not comparable, and the pseudo-$R^2$ is given either. \label{tab:reg}}
\medskip
\begin{tabular}{|l|c|c|c|c|c|}
\hline
Model  & (1) & (2) & (3) & (4) & (5) \\ 
\hline
$\log(d_{ij})$ &      -0.06** (0.03) &   -0.11** (0.05)  & -0.41*** (0.02)  & -0.35*** (0.04)  &  -0.26*** ($5\cdot 10^{-6}$) \\
$\log(T_i)$ &   &   & 0.56*** (0.01) &  0.56*** (0.01) & 0.79*** ($2\cdot 10^{-6}$) \\
$\log(T_j)$ &     &   & 0.39*** (0.01) &  0.39*** (0.01) & 0.66***  ($1.5\cdot 10^{-6}$) \\
$\log(s_{ij})$ &     &   &  &  0.19*** (0.07) & 0.53*** ($9\cdot 10^{-6}$)  \\
Countries &    &  28.3\% &   &  26.5\% & 100\% \\
\hline
$R^2$ &       0.00059   &  0.17 & 0.21 &  0.31  &  0.56 \\
MSE & 7.75 & 6.84 & 6.10 & 5.33 & 8.72 \\
AIC &        44304   &  43578  &  42131  & 41917 &   \\
\hline
\end{tabular}
\end{adjustwidth}
\end{table}
%%%%%%%%%%%%%

%\todo{which model is what is not clear between the text and the table / econometric conventions}

%\section*{Results}

\section*{Model exploration and calibration}

@@

\subsection*{Synthetic city systems}

@@

\subsection*{Application: impact of economic shock}

Based on calibrated parameter values, we design scenarios to evaluate the impact of economic shocks. These can be the consequence of diverse factors, including socio-economic, geopolitical factors (example of Brexit), and natural disasters such as COVID-19. We simulate a full border closure by setting $\gamma_D = 0$, and intermediate scenarios. Setting also $\gamma_G$ to 0 allows simulating inter-country lock-downs.
% run simulations
% note that time scales are in fact smaller - or assume that crisis keeping border closed is rather long

%%%%%%%%%%%%%%%
\section*{Discussion}
